---
title: "IXN 1 — Exploratory Data Analysis"
author: "ANONYMIZED"
date: "2023-09-08"
output:
  html_document:
    theme: flatly
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
always_allow_html: yes
font-family: DejaVu Sans
mainfont: DejaVu Sans
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#UTILITIES
library(Hmisc) # %nin% operator
library(tidyverse) #all the things
library(summarytools) #data quality
library(jtools) #Social Science regression utilities

#VIZ
library(kableExtra) #printing tables
library(ggformula) #regression syntax viz
library(vcd) #mosaic plots
library(vcdExtra) #mosaic plot helpers

#MODELLING
# library(easystats) #modelling helpers
# library(see)
# library(sjPlot)
# library(lme4)
# library(lmerTest) #for CIs in glmer 

options(readr.show_col_types = FALSE) #don't show coltypes on read_csv

```

```{r IMPORT}  

#IMPORT (wrangled) data 
df_raw <- read_csv("data/CLEAN_UtterancesForAnalysis.csv") 

#WRANGLE into DF of coded utterances 
#NOT unique utterances, 1 obs for each utterance+detail-code
df_coded <- df_raw %>% 
  #rename and factorize cols
  mutate(
    #UNIQUE IDS
    sid = factor(SID), #unique ID for utterance+detail-code
    pid = factor(PID, levels = c( #define level order so happiness first
      #HAPPINESS-FIRST    
      "bjs827ee1u", "3r2sh20ei", "4728sjuiz","7ACC0B75","92ghd48xe","iurmer28", "s294hoei",
      #SPACE-FIRST    
      "j2719eertu2","lkin27js09b","li832lin23","7382kwtue","E1D39056","8v892iige")),   
    #create unique ID for utterances
    uid = factor(as.numeric(factor(paste(pid,factor(Utterance))))), #construct a unique ID for utterances
    #recode lower case and order based on true task order
    TASK = factor(recode(Condition, "Static"="static", "Interactive"="ixn" )),
    TASK = factor(TASK, levels = c("static", "ixn")), #reorder factor levels
    #rename Notebook as DATASET
    DATASET = factor(recode(Notebook, "Happiness"="happiness", "Space"="space")),
    #create temp dataset order var
    data_order = factor(paste(TASK,"_",DATASET)), #create an order var 
    data_order = recode(data_order, "ixn _ happiness"="space-first",
                                    "ixn _ space"="happiness-first",
                                    "static _ happiness"="happiness-first",
                                    "static _ space"="space-first"),
    utterance = Utterance,
    reps_group = factor(`group`),
    reps_all = factor(`All representations`),
    #rename flags
    flag_story = `Flag Storytelling`,
    flag_correction = `Flag Correction`,
    flag_simultaneous = `Flag Simultaneous Characterization`,
    #recode and order TOP LEVEL CODES 
    code_topic = factor(Highlevel),
    code_topic = recode(code_topic, "ANALYSIS PROCESS" = "PROCESS"),
    code_topic = factor(code_topic, levels = c("PROCESS","DATASET","VARIABLE","RELATIONSHIP")),
    code_datatype = factor(`Data Type`),
    code_detail = factor(`Utterance Type`),
    timestamp = Timestamp,
    ixn = factor(interaction_used), #was interaction used?
    PNUM = factor(PNUM, levels = c("P1", "P2", "P3", "P4", "P5", "P6", 
                                   "P7", "P8", "P9", "P10", "P11","P12","P13")),
    
    ) %>% 
  select(sid,pid,PNUM,uid,TASK,DATASET,timestamp,ixn,code_topic,code_detail,code_datatype,
         flag_story, flag_correction, flag_simultaneous, utterance, reps_group, reps_all) 

#REPLACE NA in logicals to FALSE  
df_coded$flag_story[is.na(df_coded$flag_story)] <- FALSE
df_coded$flag_correction[is.na(df_coded$flag_correction)] <- FALSE
df_coded$flag_simultaneous[is.na(df_coded$flag_simultaneous)] <- FALSE

```

**There are `r nrow(df_coded)` rows in the `df_coded` dataset, where each row represents an utterance coding (i.e. utterance + detail code).  There are `r nlevels(df_coded$uid)` _unique_ utterances. The difference indicates utterances that were _dual-coded_ (i.e. two detail-level codes). No more than two codes were applied to a single utterance.  For the purposes of analysis, dual-coded utterances will be treated as two utterances, as they have two distinct (but lexically insepeperable) units of meaning. 
